---
title: "Chap_2 Coordinates"
author: "Dr K"
date: "July 18, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
```

## Zooming In
In the video, you saw different ways of using the coordinates layer to zoom in. In this exercise, we'll compare some of the techniques again.

As usual, you'll be working with the mtcars dataset, which is already cleaned up for you (cyl and am are categorical variables). Also p, a ggplot object you coded in the previous chapter, is already available. Execute p in the console to check it out.

Instructions
100 XP
Extend p with a scale_x_continuous() with limits = c(3, 6) and expand = c(0, 0). What do you see?
Try again, this time with coord_cartesian(): Set the xlim argument equal to c(3, 6). Compare the two plots.

#### Hint

Hint
Type p + scale_x_continuous(___, ___), where you fill in ___ as instructed.
Type p + coord_cartesian(___), where you fill in ___ as instructed.

```{r}
# Basic ggplot() command, coded for you
p <- ggplot(mtcars, aes(x = wt, y = hp, col = am)) + geom_point() + geom_smooth()

# Add scale_x_continuous()
p + 
  scale_x_continuous(limits = c(3, 6), expand = c(0, 0))

# Add coord_cartesian(): the proper way to zoom in
p +
  coord_cartesian(xlim = c(3, 6))
```

## Aspect Ratio
We can set the aspect ratio of a plot with coord_fixed() or coord_equal(). Both use ratio = 1 as a default. A 1:1 aspect ratio is most appropriate when two continuous variables are on the same scale, as with the iris dataset.

All variables are measured in centimeters, so it only makes sense that one unit on the plot should be the same physical distance on each axis. This gives a more truthful depiction of the relationship between the two variables since the aspect ratio can change the angle of our smoothing line. This would give an erroneous impression of the data.

Of course the underlying linear models don't change, but our perception can be influenced by the angle drawn.

Instructions
100 XP
Complete the basic scatter plot function using the iris data frame to plot Sepal.Width onto the y aesthetic, Sepal.Length onto the x and Species onto col. You should understand all the other functions used in this plot call by now. This is saved in an object called base.plot.
Write base.plot on a new line to print it out. Examine it: the plot is drawn to the dimensions of the graphics device.
Add a coord_equal() layer to force a 1:1 aspect ratio.

#### Hint

Hint
Use x = Sepal.Length, y = Sepal.Width, col = Species inside aes().
To fix the aspect ratio of base.plot(), add coord_equal() or coord_fixed() to  base.plot.

```{r}

```

## Pie Charts
The coord_polar() function converts a planar x-y Cartesian plot to polar coordinates. This can be useful if you are producing pie charts.

We can imagine two forms for pie charts - the typical filled circle, or a colored ring.

As an example, consider the stacked bar chart shown in the viewer. Imagine that we just take the y axis on the left and bend it until it loops back on itself, while expanding the right side as we go along. We'd end up with a pie chart - it's simply a bar chart transformed onto a polar coordinate system.

Typical pie charts omit all of the non-data ink, which we'll learn about in the next chapter. Pie charts are not really better than stacked bar charts, but we'll come back to this point in the fourth chapter on best practices.

The mtcars data frame is available, with cyl converted to a factor for you.

Instructions
100 XP
Create a basic stacked bar plot. Since we have univariate data and stat_bin() requires an x aesthetic, we'll have to use a dummy variable. Set x to 1 and map cyl onto fill. Assign the bar plot to wide.bar.
Add a coord_polar() layer to wide.bar. Set the argument theta to "y". This specifies the axis which would be transformed to polar coordinates.
Repeat the code for the stacked bar plot, but this time:
Set the width argument inside the geom_bar() function to 0.1 and
Use scale_x_continuous() to set the limits argument to c(0.5,1.5)). These two steps will add empty space aroung the bar on the x axis.
Assign this plot to thin.bar.
Add a coord_polar() layer to thin.bar, as you did before. There's a ring structure instead of a pie!

#### Hint

Hint
Inside aes(), assign cyl to the fill aesthetic and 1 to the x aesthetic.
Add coord_polar(theta = "y") to convert the stacked bar charts to pie charts.

```{r}
# Create a stacked bar plot: wide.bar
wide.bar <- ggplot(mtcars, aes(x = 1, fill = cyl)) +
              geom_bar()

# Convert wide.bar to pie chart
wide.bar +
  coord_polar(theta = "y")

# Create stacked bar plot: thin.bar
thin.bar <- ggplot(mtcars, aes(x = 1, fill = cyl)) +
              geom_bar(width = .1) +
              scale_x_continuous(limits = c(0.5,1.5))

# Convert thin.bar to "ring" type pie chart
thin.bar +
  coord_polar(theta = "y")
```

