---
title: "Chap_3 Themes"
author: "teacher"
date: "July 22, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
```

## Rectangles
To understand all the arguments for the themes, you'll modify an existing plot over the next series of exercises.

Here you'll focus on the rectangles of the plotting object z that has already been created for you. If you type z in the console, you can check it out. The goal is to turn z into the plot in the viewer. Do this by following the instructions step by step.

Instructions
100 XP
Plot 1: In the theme() function added to z, set the plot.background argument to element_rect(fill = myPink). myPink is already available in the workspace for you.
Plot 2: Expand your code for Plot 1 by adding a border to the plot.background. Do this by adding 2 arguments to the element_rect() function in theme(): color and size. Set them to "black" and 3, respectively.
Plot 3: we don't want the plot panels and legend to appear as they are in Plot 2. A short cut is to remove all rectangles, as defined in the theme object no_panels, and then draw the one we way in the way we want. Copy your theme() layer from Plot 2 and add it to no_panels.

#### Hint

Hint
For the second plot, the plot.background argument inside theme() should be set to element_rect(fill = myPink, color = "black", size = 3).
To remove the unwanted backgrounds, add the predefined theme, uniform_panels to the third plot.

```{r }
#need to figure out how to put z in here
z <- ggplot(mtcars,aes(x=wt,y=mpg, col=cyl))

myPink <- c("#99000D")
# Starting point
z + geom_point()

# Plot 1: Change the plot background fill to myPink
z + 
  theme(plot.background = element_rect(fill = myPink))

# Plot 2: Adjust the border to be a black line of size 3
z + 
  theme(plot.background = element_rect(fill = myPink, color = "black", size = 3)) # expanded from plot 1

# Theme to remove all rectangles
no_panels <- theme(rect = element_blank())

# Plot 3: Combine custom themes
z +
  no_panels +
  theme(plot.background = element_rect(fill = myPink, color = "black", size = 3))  # from plot 2
```

## Lines
To change the appearance of lines use the element_line() function.

The plot you created in the last exercise, with the fancy pink background, is available as the plotting object z. Your goal is to produce the plot in the viewer - no grid lines, but red axes and tick marks.

For each of the arguments that specify lines, use element_line() to modify attributes. e.g. element_line(color = "red").

Remember, to remove a non-data element, use element_blank().

Instructions
100 XP
Starting with object z, add a theme() function to:

remove the grid lines using the panel.grid argument.
add red axis lines using the axis.line argument.
change the tick marks to red using the axis.ticks argument, similar to how you specified axis.line.

#### Hint

Hint
Set panel.grid = element_blank().
Set both axis.line and axis.ticks to element_line(color = "red").

```{r }
# Extend z with theme() function and 3 args
z + 
  theme(panel.grid = element_blank(),
        axis.line = element_line(color = "red"),
        axis.ticks = element_line(color = "red"))
```

## Text
Next we can make the text on your plot prettier and easier to spot. You can do this through the element_text() function and by passing the appropriate arguments inside the theme() function.

As before, the plot you've created in the previous exercise is available as z. The plot you should end up with after successfully completing this exercises is shown in the viewer.

Instructions
100 XP
Starting from z, add a theme() function to:

Change the appearance of the strip text, that is the text in the facet strips. Specify strip.text with element_text(). The size of the text should be 16, the color should be myRed, a color that is predefined for you.
Change the axis titles. Specify both axes with the axis.title argument and use element_text() to set the parameters: color = myRed, hjust = 0 (to put the text in the bottom left corner) and face = "italic".
Make the axis text black using the axis.text argument to do so.

#### Hint

Hint
Use element_text(size = ___, color = ___) to specify strip.text.
Use element_text(color = ___, hjust = ___, face = ___) to specify axis.title. Fill in the blanks!
Use element_text(color = ___) to specify axis.text.

```{r }
# Original plot, color provided
z
myRed <- myPink

# Extend z with theme() function and 3 args
z +
  theme(strip.text = element_text(size = 16, color = myRed),
        axis.title = element_text(color = myRed, hjust = 0, face = "italic"),
        axis.text = element_text(color = "black"))
```

## Legends
The themes layer also allows you to specify the appearance and location of legends.

The plot you've coded up to now is available as z. It's also displayed in the viewer. Solve the instructions and compare the resulting plots with the plot you started with.

Instructions
100 XP
Add a theme() function to z to change the legend's location. Do this by specifying the legend.position argument to be c(0.85, 0.85). This will make the legend appear in the top right of the plot, inside the third facet.
Instead of a vertical list of legend entries, you might want to have the different entries next to each other. Starting from z, add a theme() function in which you specify legend.direction to be "horizontal".
You can also change the locations of legends by name: set legend.position to "bottom".
Finally, you can remove the legend entirely, by setting legend.position to "none".

#### Hint

Hint
Follow the instructions closely. Each time, you should simply add one theme() function with a single argument to z. The instructions are not incremental here.

```{r}
# Move legend by position
z +
  theme(legend.position = c(0.85, 0.85))

# Change direction
z +
  theme(legend.direction = "horizontal")

# Change location by name
z +
  theme(legend.position = "bottom")

# Remove legend entirely
z +
  theme(legend.position = "none")
```
## Rectangles
To understand all the arguments for the themes, you'll modify an existing plot over the next series of exercises.

Here you'll focus on the rectangles of the plotting object z that has already been created for you. If you type z in the console, you can check it out. The goal is to turn z into the plot in the viewer. Do this by following the instructions step by step.

Instructions
100 XP
Plot 1: In the theme() function added to z, set the plot.background argument to element_rect(fill = myPink). myPink is already available in the workspace for you.
Plot 2: Expand your code for Plot 1 by adding a border to the plot.background. Do this by adding 2 arguments to the element_rect() function in theme(): color and size. Set them to "black" and 3, respectively.
Plot 3: we don't want the plot panels and legend to appear as they are in Plot 2. A short cut is to remove all rectangles, as defined in the theme object no_panels, and then draw the one we way in the way we want. Copy your theme() layer from Plot 2 and add it to no_panels.

#### Hint

Hint
For the second plot, the plot.background argument inside theme() should be set to element_rect(fill = myPink, color = "black", size = 3).
To remove the unwanted backgrounds, add the predefined theme, uniform_panels to the third plot.

```{r}
# Starting point
z

# Plot 1: Change the plot background fill to myPink
z + 
  theme(plot.background = element_rect(fill = myPink))

# Plot 2: Adjust the border to be a black line of size 3
z + 
  theme(plot.background = element_rect(fill = myPink, color = "black", size = 3)) # expanded from plot 1

# Theme to remove all rectangles
no_panels <- theme(rect = element_blank())

# Plot 3: Combine custom themes
z +
  no_panels +
  theme(plot.background = element_rect(fill = myPink, color = "black", size = 3))  # from plot 2
```
## Lines
To change the appearance of lines use the element_line() function.

The plot you created in the last exercise, with the fancy pink background, is available as the plotting object z. Your goal is to produce the plot in the viewer - no grid lines, but red axes and tick marks.

For each of the arguments that specify lines, use element_line() to modify attributes. e.g. element_line(color = "red").

Remember, to remove a non-data element, use element_blank().

Instructions
100 XP
Starting with object z, add a theme() function to:

remove the grid lines using the panel.grid argument.
add red axis lines using the axis.line argument.
change the tick marks to red using the axis.ticks argument, similar to how you specified axis.line.

#### Hint

Hint
Set panel.grid = element_blank().
Set both axis.line and axis.ticks to element_line(color = "red").

```{r}
# Extend z with theme() function and 3 args
z + 
  theme(panel.grid = element_blank(),
        axis.line = element_line(color = "red"),
        axis.ticks = element_line(color = "red"))
```

## Text
Next we can make the text on your plot prettier and easier to spot. You can do this through the element_text() function and by passing the appropriate arguments inside the theme() function.

As before, the plot you've created in the previous exercise is available as z. The plot you should end up with after successfully completing this exercises is shown in the viewer.

Instructions
100 XP
Starting from z, add a theme() function to:

Change the appearance of the strip text, that is the text in the facet strips. Specify strip.text with element_text(). The size of the text should be 16, the color should be myRed, a color that is predefined for you.
Change the axis titles. Specify both axes with the axis.title argument and use element_text() to set the parameters: color = myRed, hjust = 0 (to put the text in the bottom left corner) and face = "italic".
Make the axis text black using the axis.text argument to do so.

#### Hint

Hint
Use element_text(size = ___, color = ___) to specify strip.text.
Use element_text(color = ___, hjust = ___, face = ___) to specify axis.title. Fill in the blanks!
Use element_text(color = ___) to specify axis.text.

```{r}
# Original plot, color provided
z
myRed

# Extend z with theme() function and 3 args
z +
  theme(strip.text = element_text(size = 16, color = myRed),
        axis.title = element_text(color = myRed, hjust = 0, face = "italic"),
        axis.text = element_text(color = "black"))
```
## Legends
The themes layer also allows you to specify the appearance and location of legends.

The plot you've coded up to now is available as z. It's also displayed in the viewer. Solve the instructions and compare the resulting plots with the plot you started with.

Instructions
100 XP
Add a theme() function to z to change the legend's location. Do this by specifying the legend.position argument to be c(0.85, 0.85). This will make the legend appear in the top right of the plot, inside the third facet.
Instead of a vertical list of legend entries, you might want to have the different entries next to each other. Starting from z, add a theme() function in which you specify legend.direction to be "horizontal".
You can also change the locations of legends by name: set legend.position to "bottom".
Finally, you can remove the legend entirely, by setting legend.position to "none".

#### Hint

Hint
Follow the instructions closely. Each time, you should simply add one theme() function with a single argument to z. The instructions are not incremental here.

```{r}
# Move legend by position
z +
  theme(legend.position = c(0.85, 0.85))

# Change direction
z +
  theme(legend.direction = "horizontal")

# Change location by name
z +
  theme(legend.position = "bottom")

# Remove legend entirely
z +
  theme(legend.position = "none")
```
## Positions
The different rectangles of your plot have spacing between them. There's spacing between the facets, between the axis labels and the plot rectangle, between the plot rectangle and the entire panel background, etc. Let's experiment!

The last plot you created in the previous exercise, without a legend, is available as z.

Instructions
100 XP
Suppose you want to have more spacing between the different facets. You can control this by specifying panel.spacing.x inside a theme() function you add to z. For the argument value, you should pass a unit object. To achieve this, load the grid package with library(). Next, set panel.spacing.x to unit(2, "cm").
Copy, adapt and paste the plot command for the previous instruction: to adjust the plot margin, set plot.margin to unit(c(1,2,1,1), "cm") (spacing for top, right, bottom, and left margins).

#### Hint

Hint
For this exercise, you need panel.spacing.x = ___ and plot.margin = ___ inside a theme() function you add to z.

```{r}
# Increase spacing between facets
library(grid)
z + theme(panel.spacing.x = unit(2, "cm"))

# Adjust the plot margin
z + theme(panel.spacing.x = unit(2, "cm"),
          plot.margin = unit(c(1,2,1,1), "cm"))
```

## Updating Themes
Building your themes every time from scratch can become a pain and unnecessarily bloat your scripts. In the following exercises, we'll practice different ways of managing, updating and saving themes.

A plot object z2 is already created for you on the right. It shows mpg against wt for the mtcars dataset, faceted according to cyl. Also the colors myPink and myRed are available. In the previous exercises you've already customized the rectangles, lines and text on the plot. This theme layer is now separately stored as theme_pink, as shown in the sample code.

theme_update() updates the default theme used by ggplot2. The arguments for theme_update() are the same as for theme(). When you call theme_update() and assign it to an object (e.g. called old), that object stores the current default theme, and the arguments update the default theme. If you want to restore the previous default theme, you can get it back by using theme_update() again. Let's see how:

Instructions
100 XP
1 - "Apply" theme_pink to z2 to carry out all customizations.
2 - Instead of applying theme_pink, use theme_update(). This function returns an object that contains the previous theme settings, so that you can restore it later. Assign the output of theme_update() to an object called old.
3 - Plot z2 again, after the theme_update() call. The resulting plot has the same appearance as the previous one - but now you don't need to call theme() explictly.
4 - Restore the old theme using theme_set(old) and plot z2 again. It's back to the original default theme.

#### Hint

Hint
You don't have to edit the theme, theme_pink. To apply it to z2, just type z2 + theme_pink. It works like magic!
Don't change the theme_update() command, just assign it's result to old. Like this: old <- ....
Don't forget to write out z2 to plot it the second time.
Restore the old theme using the command that is given in the instructions.

```{r eval=FALSE}
# Original plot
z2

# Theme layer saved as an object, theme_pink
theme_pink <- theme(panel.background = element_blank(),
                    legend.key = element_blank(),
                    legend.background = element_blank(),
                    strip.background = element_blank(),
                    plot.background = element_rect(fill = myPink, color = "black", size = 3),
                    panel.grid = element_blank(),
                    axis.line = element_line(color = "red"),
                    axis.ticks = element_line(color = "red"),
                    strip.text = element_text(size = 16, color = myRed),
                    axis.title.y = element_text(color = myRed, hjust = 0, face = "italic"),
                    axis.title.x = element_text(color = myRed, hjust = 0, face = "italic"),
                    axis.text = element_text(color = "black"),
                    legend.position = "none")
  
# 1 - Apply theme_pink to z2
z2 + 
  theme_pink

# 2 - Update the default theme, and at the same time
# assign the old theme to the object old.
old <- theme_update(panel.background = element_blank(),
             legend.key = element_blank(),
             legend.background = element_blank(),
             strip.background = element_blank(),
             plot.background = element_rect(fill = myPink, color = "black", size = 3),
             panel.grid = element_blank(),
             axis.line = element_line(color = "red"),
             axis.ticks = element_line(color = "red"),
             strip.text = element_text(size = 16, color = myRed),
             axis.title.y = element_text(color = myRed, hjust = 0, face = "italic"),
             axis.title.x = element_text(color = myRed, hjust = 0, face = "italic"),
             axis.text = element_text(color = "black"),
             legend.position = "none")

# 3 - Display the plot z2 - new default theme used
z2

# 4 - Restore the old default theme
theme_set(old)

# Display the plot z2 - old theme restored
z2
```
## Exploring ggthemes
There are many themes available by default in ggplot2: theme_bw(), theme_classic(), theme_gray(), etc. In the previous exercise, you saw that you can apply these themes to all following plots, with theme_set():

theme_set(theme_bw())
But you can also apply them on an individual plot, with:

... + theme_bw()
You can also extend these themes with your own modifications. In this exercise, you'll experiment with this and use some preset templates available from the ggthemes package. The workspace already contains the same basic plot from before under the name z2.

Instructions
100 XP
Instructions
100 XP
Create a custom theme, assigning it to custom_theme.
Call theme_tufte() with no arguments.
Add a call to theme() as follows.
Set legend.position to c(0.9, 0.9).
Set legend.title to an "italic" text of size 12. Use element_text(face = ___, size = ___).
Set axis.title to a "bold" text of size 14. Use element_text(face = ___, size = ___).
Plot z2 with the customized theme. (You don't need parentheses.)
Make custom_theme the default by calling theme_set().
Plot z2 again.

#### Hint

Hint
To apply theme_tufte() to a specific plot, use + theme_tufte().
Then add theme(legend.position = ___, legend.title = ___, axis.title = ___). Replace the ___ with the instructed arguments.
a_plot + a_theme uses that theme.
theme_set(a_theme) sets the theme.

```{r eval=FALSE}
# Original plot
z2

# Load ggthemes
library(ggthemes)

# Apply theme_tufte, plot additional modifications
custom_theme <- theme_tufte() +
  theme(legend.position = c(0.9, 0.9),
        legend.title = element_text(face = "italic", size = 12),
        axis.title = element_text(face = "bold", size = 14))

# Draw the customized plot
z2 + custom_theme
 
# Use theme set to set custom theme as default
theme_set(custom_theme)

# Plot z2 again
z2
```
## Bar Plots (1)
In the video we saw why "dynamite plots" (bar plots with error bars) are not well suited for their intended purpose of depicting distributions. If you really want error bars on bar plots, you can still get that. However, you'll need to set the positions manually. A point geom will typically serve you much better.

We saw an example of a dynamite plot earlier in this course. Let's return to that code and make sure you know how to handle it. We'll use the mtcars dataset for examples. The first part of this exercise will just be a refresher, then we'll get into some details.

Instructions
100 XP
To begin with, create the base layers where you map wt on the y axis and cyl on the x axis. cyl and am have already been converted to a factor variable for you. Store these base layers as an object m.
Draw a dynamite plot filling in the details for the bars in the first stat_summary() function. Set fun.y to mean, geom to "bar" and give the bars a "skyblue" fill. The stat_summary() function that adds error bars is coded for you.

#### Hint

Hint
Only specify the x and y mappings for the first plot, as indicated in the sample code.
For the second instruction, fill in the parts in the sample code that are identified with ___ as instructed.

```{r}
# Base layers
m <- ggplot(mtcars, aes(x = cyl, y = wt))

# Draw dynamite plot
m +
  stat_summary(fun.y = mean, geom = "bar", fill = "skyblue") +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), geom = "errorbar", width = 0.1)
```

## Bar Plots (2)
In the previous exercise we used the mtcars dataset to draw a dynamite plot about the weight of the cars per cylinder type.

In this exercise we will add a distinction between transmission type, am, for the dynamite plots.

Instructions
100 XP
Update m so that we split the bars according to transmission type, am. Note that for bar plots, we want to change the col as well as the fill.
Plot 1 is already coded for you, but it is not optimal. Let's fix that in the following instructions.
Plot 2: copy the code for Plot 1 and set the position to "dodge" - this also doesn't work, because the default dodging is different for the different stat_summary() functions.
Plot 3: copy the code for Plot 2 and set the position to the object posn.d, which defines a dodge position using position_dodge(0.9).

#### Hint

Hint
When creating the base layers, col and fill have to be set to the same variable, am.
Plot 1 is already coded for you, so you don't have to change anything here.
In Plot 2, you should set position = "dodge" for both stat_summary() layers.
In Plot 3, set the position to the earlier created posn.d. Do this for both stat_summary() layers. Watch out, no quotes this time!

```{r}
# Base layers
m <- ggplot(mtcars, aes(x = cyl,y = wt, col = am, fill = am))

# Plot 1: Draw dynamite plot
m +
  stat_summary(fun.y = mean, geom = "bar") +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), geom = "errorbar", width = 0.1)

# Plot 2: Set position dodge in each stat function
m +
  stat_summary(fun.y = mean, geom = "bar", position = "dodge") +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "errorbar", width = 0.1, position = "dodge")

# Set your dodge posn manually
posn.d <- position_dodge(0.9)

# Plot 3: Redraw dynamite plot
m +
  stat_summary(fun.y = mean, geom = "bar", position = posn.d) +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), geom = "errorbar", width = 0.1, position = posn.d)
```

## Bar Plots (3)
If it is appropriate to use bar plots (see the video for a discussion!), then it would also be nice to give an impression of the number of values in each group.

stat_summary() doesn't keep track of the count. stat_sum() does (that's the whole point), but it's difficult to access. In this case, the most straightforward thing to do is calculate exactly what we want to plot beforehand. For this exercise we've created a summary data frame called mtcars.cyl which contains the average (wt.avg), standard deviations (sd) and count (n) of car weights, according to cylinders, cyl. It also contains the proportion (prop) of each cylinder represented in the entire dataset. Use the console to familiarize yourself with the mtcars.cyl data frame.

Instructions
100 XP
Instructions
100 XP
Establish the base layers. Use the mtcars.cyl dataset and map cyl onto x and wt.avg onto y. Call the resulting ggplot object m.
Plot 1: Starting from m
Add a geom_bar() layer
In this geom_bar(), set the attribute stat to "identity" and the attribute fill to "skyblue".
Plot 2: geom_col() is a shortcut for geom_bar(stat = "identity"), for when your data already has counts.
Add a geom_col() layer. This time just specify the fill argument as above. The result should be the same, but here we don't have to specify the stat argument.
Plot 3: Starting from Plot 2,
Add the argument width = mtcars.cyl$prop inside the geom_col() layer. mtcars.cyl$prop is a column that represents the proportion of each group.
Plot 4: Starting from Plot 3,
Add error bars to create a dynamite plot using geom_errorbar()
Inside aes() of this geom_errorbar() layer, specify: ymin = wt.avg - sd and ymax = wt.avg + sd.
Inside geom_errorbar(), but outside its aes(), set the width = 0.1 to control the width of the error bars.

#### Hint

Hint
For the base layers, use aes(x = cyl, y = wt.avg).
For Plot 1, specify stat = "identity", fill = "skyblue" inside geom_bar().
For Plot 2, specify fill = "skyblue" inside geom_col().
For Plot 3, start from Plot 2 and add a width = mtcars.cyl$prop argument inside geom_col().
For Plot 4, start from Plot 3 and add geom_errorbar(aes(ymin = ___, ymax = ___), width = 0.1). Fill in the ___ parts.

```{r eval=FALSE}
# Base layers
m <- ggplot(mtcars.cyl, aes(x = cyl, y = wt.avg))

# Plot 1: Draw bar plot with geom_bar
m + geom_bar(stat = "identity", fill = "skyblue")

# Plot 2: Draw bar plot with geom_col
m + geom_col(fill = "skyblue")

# Plot 3: geom_col with variable widths.
m + geom_col(fill = "skyblue", width = mtcars.cyl$prop)

# Plot 4: Add error bars
m +
  geom_col(fill = "skyblue", width = mtcars.cyl$prop) +
  geom_errorbar(aes(ymin = wt.avg - sd, ymax = wt.avg + sd), width = 0.1)
```

## Pie Charts (1)
In this example we're going to consider a typical use of pie charts - a categorical variable as the proportion of another categorical variable. For example, the proportion of each transmission type am, in each cylinder, cyl class.

The first plotting function in the editor should be familiar to you by now. It's a straightforward bar chart with position = "fill", as shown in the viewer. This is already a good solution to the problem at hand! Let's take it one step further and convert this plot in a pie chart.

Instructions
100 XP
Instructions
100 XP
Adapt the code for the bar chart in the editor to turn it into a good looking pie chart:

Transform the bar plot into a facetted plot: add a facet_grid() call to split columns by cyl. Remember to use formula notation here ROW ~ COL.

For the moment, each facet will only have one category because cyl is also mapped onto x. Use a dummy aesthetic for the x. Change the aes() function such that factor(1) maps onto x.

Add a coord_polar() call where you specify the theta to "y".

This is already pretty good, but to remove all non-data ink add a theme_void() layer. (This is the first time we've seen this theme, but you should be familiar with themes already).

There's a small hole in the center of the pies. Inside geom_bar() set width = 1 so that the bars fill up the entire width resulting in a full pie chart.

#### Hint

Hint
The formula within facet_grid() should be . ~ cyl.
To use the dummy aesthetic appropriately change x = cyl to x = factor(1).
To transform the axis, use coord_polar(theta = "y").
To remove all non-data ink use theme_void().
To fix the hole in the pie charts' centers, update geom_bar(): geom_bar(position = "fill", width = 1).

```{r}
# Bar chart
ggplot(mtcars, aes(x = cyl, fill = am)) +
  geom_bar(position = "fill")

# Convert bar chart to pie chart
ggplot(mtcars, aes(x = factor(1), fill = am)) +
  geom_bar(position = "fill", width = 1) +
  facet_grid(. ~ cyl) + 
  coord_polar(theta = "y") +
  theme_void()
```

## Pie Charts (2)
In the previous example, we looked at one categorical variable (am) as a proportion of another (cyl). Here, we're interested in two or more categorical variables, independent of each other. The many pie charts in the viewer is an unsatisfactory visualization. We're interested in the relationship between all these variables (e.g. where are 8 cylinder cars represented on the Transmission, Gear and Carburetor variables?) Perhaps we also want continuous variables, such as weight. How can we combine all this information?

The trick is to use a parallel coordinates plot, like this one. Each variable is plotted on its own parallel axis. Individual observations are connected with lines, colored according to a variable of interest. This is a surprisingly useful visualization since we can combine many variables, even if they are on entirely different scales.

A word of caution though: typically it is very taboo to draw lines in this way. It's the reason why we don't draw lines across levels of a nominal variable - the order, and thus the slope of the line, is meaningless. Parallel plots are a (very useful) exception to the rule!

Instructions
100 XP
am is variable 9 in the mtcars data frame. Assign this number to group_by_am. The object my_names_am will contain a numeric vector from 1 - 11 excluding the column with am. These will be our parallel axes.
Fill in the ggparcoord() function.
The first argument is the data frame you're using. mtcars in our case.
The second argument is the number of the columns to plot (use my_names_am),
groupColumn specifies the column number of the grouping variable (use group_by_am)
alpha, the opacity, should be set to 0.8

#### Hint

Hint
Set group_by_am to 9 for the first instruction.
For the second instruction, use ggparcoord(mtcars, my_names_am, groupColumn = group_by_am, alpha = 0.8)

```{r}
# Parallel coordinates plot using GGally
library(GGally)

# All columns except am
group_by_am <- 9
my_names_am <- (1:11)[-group_by_am]

# Basic parallel plot - each variable plotted as a z-score transformation
ggparcoord(mtcars, my_names_am, groupColumn = group_by_am, alpha = 0.8)
```

## Plot Matrix (1)
The parallel coordinate plot from the last exercise is an excellent example of an exploratory plot. It presents a massive amount of information and allows the specialist to explore many relationships all at once. Another great example is a plot matrix (a SPLOM, from scatter plot matrix).

GGally::ggpairs(mtcars2) will produce the plot of a selection of the mtcars dataset, mtcars2, in the viewer. Depending on the nature of the dataset a specific plot type will be produced and if both variables are continuous the correlation (rho) will also be calculated.

The relationship between the variables drat and mpg is shown in two areas. What is the correlation between these two variables?

Instructions
50 XP
Possible Answers
0.659
0.681
0.888
0.419
This value is not given.

#### Hint

Hint
Run the given command in the console. What does the error message say?0

## Plot Matrix (2)
On the right, another plot matrix is shown based off of a subset of the mtcars data frame. What is the relationship between disp and cyl?

Instructions
50 XP
Possible Answers
As cyl increases, the distributuion of disp stays the same.
As cyl increases, disp decreases.
As cyl increases, so to does the disp.
As cyl increases, the variance in disp decreases.
This information is not given.

#### Hint

Hint
Look for disp on the y axis and go to the second plot, for cyl, in this row.

## Heat Maps
In the video you saw reasons for not using heat maps. Nonetheless, you may encounter a case in which you really do want to use one. Luckily, they're fairly straightforward to produce in ggplot2.

We begin by specifying two categorical variables for the x and y aesthetics. At the intersection of each category we'll draw a box, except here we call it a tile, using the geom_tile() layer. Then we will fill each tile with a continuous variable.

We'll produce the heat map we saw in the video with the built-in barley dataset. The barley dataset is in the lattice package and has already been loaded for you. Begin by exploring the structure of the data in the console using str().

Instructions
100 XP
Instructions
100 XP
Reproduce the heat map shown in the viewer in different steps:

Define the data and the aesthetics layer. Using the barley dataset, map year onto x, variety onto y and fill according to yield
Add a geom_tile() to build the heat maps.
So far the entire dataset is plotted on one heat map. Add a facet_wrap() function to get a facetted plot. Use the formula ~ site (without the dot!) and set ncol = 1. By default, the names of the farms will be above the panels, not to the side (as we get with facet_grid()).
brewer.pal() from the RColorBrewer package has been used to create a "Reds" color palette. The hexadecimal color codes are stored in the myColors object. Add the scale_fill_gradientn() function and specify the colors argument correctly to give the heat maps a reddish look.

#### Hint

Hint
Add a geom_tile() layer.
For facets, use facet_wrap( ~ site, ncol = 1), not facet_grid(. ~ site).
The scale function should produce an n gradient of colors (which is myColors). Pay attention to the distinction between color and fill. Therefore, use scale_fill_gradientn(colors = myColors).

```{r}
require(RColorBrewer)
require(lattice)
# Create color palette
myColors <- brewer.pal(9, "Reds")

# Build the heat map from scratch
ggplot(barley, aes(x = year, y = variety, fill = yield)) +
  geom_tile() + # Geom layer
  facet_wrap( ~ site, ncol = 1) + # Facet layer
  scale_fill_gradientn(colors = myColors) # Adjust colors
```

## Heat Maps Alternatives (1)
There are several alternatives to heat maps. The best choice really depends on the data and the story you want to tell with this data. If there is a time component, the most obvious choice is a line plot like what we see in the viewer. Can you come up with the correct commands to create a similar looking plot?

The barley dataset is already available in the workspace. Feel free to check out its structure before you start!

Instructions
100 XP
The line plot might be a good alternative:
Base layer: same dataset, map year onto x, yield onto y and variety onto col as well as onto group!
Add the appropriate geom for this line plot; no additional arguments are needed.
Add facetting with the same formula as in the heat map plot, instead of ncol, set nrow to 1.

#### Hint

Hint
In the aes() layer of the ggplot() command, you should have 4 aesthetics: x, y, col and group. col and group should both be set to variety.
geom_line() is what you're looking for!
Use facet_wrap() with ~ site. Don't forget to use nrow, not ncol.

```{r}
# The heat map we want to replace
# Don't remove, it's here to help you!
myColors <- brewer.pal(9, "Reds")
ggplot(barley, aes(x = year, y = variety, fill = yield)) +
  geom_tile() +
  facet_wrap( ~ site, ncol = 1) +
  scale_fill_gradientn(colors = myColors)

# Line plot; set the aes, geom and facet
ggplot(barley, aes(x = year, y = yield, color = variety, group = variety)) + 
  geom_line() +
  facet_wrap( ~ site, nrow = 1)
```

## Heat Maps Alternatives (2)
In the videos we saw two methods for depicting overlapping measurements of spread. You can use dodged error bars or you can use overlapping transparent ribbons (shown in the viewer). In this exercise we'll try to recreate the second option, the transparent ribbons.

The barley dataset is available. You can use str(barley) to refresh its structure before heading over to the instructions.

Instructions
100 XP
Create a plot, similar to the one in the viewer, from scratch by following these steps:

Base layer: use the barley dataset. Try to come up with the correct mappings for x, y, col, group and fill.
Add a stat_summary() function for the mean. Specify fun.y to be mean and set geom to "line".
Add a stat_summary() function for the ribbons. Set fun.data = mean_sdl and fun.args = list(mult = 1) to have a ribbon that spans over one standard deviation in both directions. Use geom = "ribbon" and set col = NA and alpha = 0.1.

#### Hint
Hint
For the base layer, map site onto col, group and fill!
stat_summary(fun.y = mean, geom = "line") will add the line to the plot. Although mean_sdl() is a function, inside stat_summary() we don't need to use (); just call it by name: mean_sdl.
stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1)) is a good start, but you have to specify three more arguments to get the plot you want: geom = "ribbon", alpha = 0.1 and col = NA.

```{r}
# Create overlapping ribbon plot from scratch
ggplot(barley, aes(x = year, y = yield, col = site, group = site, fill = site)) +
  stat_summary(fun.y = mean, geom = "line") +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), geom = "ribbon", alpha = 0.1, col = NA)
```
## 
